### 2020-09-27 doc_arXiv/doc_arXiv.Rmd
use train but not test outcomes

---
title: Improving Outcome Predictions for Patients Receiving Mechanical Circulatory Support by Optimizing Imputation of Missing Values
authors:
  - name: Byron C. Jaeger
    thanks: Source code available at https://github.com/bcjaeger/INTERMACS-missing-data
    department: Department of Biostatistics
    affiliation: University of Alabama at Birmingham
    location: Birmingham, AL 35211
    email: bcjaeger@uab.edu
abstract: |
  \textbf{Background} Risk predictions play an important role in clinical decision making. When developing risk prediction models, practitioners often impute missing values to the mean. The purpose of this article is to evaluate the impact of applying different strategies to impute missing values on the prognostic accuracy of prediction models fitted to the imputed data. A secondary objective was to compare the accuracy of different imputation methods. To complete these objectives, we used data from the Interagency Registry for Mechanically Assisted Circulatory Support (INTERMACS).  \newline\textbf{Methods and Results} We applied all pairwise combinations of <fill in> different strategies to impute missing values and three different strategies to fit a risk prediction model for mortality and transplant after receiving mechanical circulatory support. Model performance was compared using Concordance (i.e. C-index), Brier Score, and Net reclassification index. Results indicated that multiple  imputation consistently resulted in  superior prognostic models compared  to  other missing data strategies, particularly imputation to the mean. <fill in results> \newline\textbf{Conclusion} Selecting an optimal strategy to handle missing values can have a substantial impact on improving model accuracy.  In the current analysis, multiple imputation emerged as an optimal strategy to handle missing values in the INTERMACS data. The evaluation and selection of the optimal missing data strategy in this work has the potential to improve risk predictions for other longitudinal registries.
keywords:
  - Missing Data, 
  - INTERMACS, 
  - imputation, 
  - heart failure, 
  - mortality, 
  - risk prediction
bibliography: references.bib
output: rticles::arxiv_article
header-includes:
   - \usepackage{placeins}
   - \newcommand{\ie}{\textit{i.e., }}
   - \newcommand{\eg}{\textit{e.g., }}
   - \newcommand{\cstat}{\widehat{\textrm{C}}(t)}
   - \newcommand{\bstat}{\widehat{\textrm{BS}}(t)}
   - \newcommand{\ibstat}{\mathcal{\widehat{BS}}(\tau)}
---

# Introduction
\label{sec:introduction}

Heart disease is a leading cause of death in the United States. Heart failure, a primary component of heart disease, affects over 6 million Americans, and for ~10\% of these patients medical management is no longer effective \cite{benjamin2017heart,national2017health}. Mechanical circulatory support (MCS) is a surgical intervention in which a mechanical device is implanted in parallel to the heart to improve circulation \cite{patel2014contemporary}. Typically, MCS is used while a patient waits for a heart transplant (bridge-to-transplant) or in some cases as an alternative to transplant (destination therapy) \cite{slaughter2009advanced}. Over 250,000 patients could benefit from MCS \cite{miller2011left}. However, less than 4,000 new patients receive a long-term MCS device each year, with widely heterogeneous outcomes \cite{stewart2011keeping}. The 2-year survival on MCS ranges from 61\% for destination therapy to 78\% for bridge-to-transplant \cite{patel2014contemporary}. Therefore, there is great need for reliable predictions of patient-specific risk to experience adverse events after receiving MCS. This information can be used to improve patient selection for MCS, inform the design of next generation pumps, and refine patient management strategies.

The Interagency Registry  for  Mechanically  Assisted  Circulatory  Support (INTERMACS) was launched to improve MCS patient outcomes through the collection of patient characteristics, medical events, and long terms outcomes at a nation wide level. Currently, INTERMACS comprises over 20,000 patients who have received a mechanically assisted circulatory support device (MCSD). As the largest registry for data on patients receiving MCSD, INTERMACS has been leveraged to develop numerous risk prediction models for mortality and other types of adverse events that may occur after receiving a device \cite{kirklin2017eighth, kormos2019society, Adamo950}. A critical step for developing risk prediction models is handling missing data, which is common in a “real world” database that is not designed for the rigor of data  completeness found in a  clinical trial. Several variables in the INTERMACS data have high rates of missing values. While some of the missing data is related to data entry error, other missing values may  relate  to instability of a patient’s circulatory status at the time of implant.  For instance, hemodynamic laboratory values are missing for patients in whom  invasive catheterizations were not performed. Patients without invasive cathaderizations  may be too ill (unstable) to tolerate the procedure. Instead, such patients may, after determination of basic hemodynamics, proceed directly to device placement.

Multiple strategies have been applied to handle missing values in the INTERMACS registry. Typically, missing values have been imputed to the mean \cite{kormos2019society}. Loghmanpour et al. combined several imputed datasets from random forest models into a single imputed dataset. Grady et al. handled missing data by assigning values based on the clinical state of the patient (physical components of heath related quality of life scores set to zero if the patient was too sick to complete the instrument) \cite{grady2015change}. This article extends previous work by formally quantifying the effect of differing strategies to handle missing data for the INTERMACS registry on downstream prediction models that are fitted to the imputed data.