### 2020-11-01 doc_arXiv/doc_arXiv.tex

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}

\title{Improving Outcome Predictions for Patients Receiving Mechanical
Circulatory Support by Optimizing Imputation of Missing Values}

\author{
    Byron C. Jaeger
    \thanks{Source code available at
\url{https://github.com/bcjaeger/INTERMACS-missing-data}}
   \\
    Department of Biostatistics \\
    University of Alabama at Birmingham \\
  Birmingham, AL 35211 \\
  \texttt{\href{mailto:bcjaeger@uab.edu}{\nolinkurl{bcjaeger@uab.edu}}} \\
  }

\begin{document}
\maketitle

\def\tightlist{}

\begin{abstract}
\textbf{Background} Risk predictions play an important role in clinical
decision making. When developing risk prediction models, practitioners
often impute missing values to the mean. The purpose of this article is
to evaluate the impact of applying different strategies to impute
missing values on the prognostic accuracy of prediction models fitted to
the imputed data. A secondary objective was to compare the accuracy of
different imputation methods. To complete these objectives, we used data
from the Interagency Registry for Mechanically Assisted Circulatory
Support (INTERMACS). \newline\textbf{Methods and Results} We applied
thirteen different strategies to impute missing values in combination
with three different strategies to fit a risk prediction model for
mortality and transplant after receiving mechanical circulatory support.
Model performance was measured by 12-month discrimination, calibration
and net reclassification index. Results indicated that multiple
imputation consistently provided prediction models with greater
prognostic accuracy than other missing data strategies, particularly
imputation to the mean. \newline\textbf{Conclusion} Selecting an optimal
strategy to handle missing values impacts the prognostic accuracy of
downstream models. In the current analysis, multiple imputation emerged
as an optimal strategy to handle missing values in the INTERMACS data.
The current study shows that evaluation and selection of an optimal
strategy to impute missing data has the potential to improve prognostic
accuracy of risk predictions for other longitudinal registries.
\end{abstract}

\keywords{
    Missing Data,
   \and
    INTERMACS,
   \and
    imputation,
   \and
    heart failure,
   \and
    mortality,
   \and
    risk prediction
  }

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\label{sec:introduction}

\linenumbers
\doublespacing

Heart disease is a leading cause of death in the United States. Heart
failure, a primary component of heart disease, affects over 6 million
Americans, and for \textasciitilde10\% of these patients medical
management is no longer effective
\cite{benjamin2017heart,national2017health}. Mechanical circulatory
support (MCS) is a surgical intervention in which a mechanical device is
implanted in parallel to the heart to improve circulation
\cite{patel2014contemporary}. Typically, MCS is used while a patient
waits for a heart transplant (bridge-to-transplant) or in some cases as
an alternative to transplant (destination therapy)
\cite{slaughter2009advanced}. Over 250,000 patients could benefit from
MCS \cite{miller2011left}. However, less than 4,000 new patients receive
a long-term MCS device each year, with widely heterogeneous outcomes
\cite{stewart2011keeping}. The 2-year survival on MCS ranges from 61\%
for destination therapy to 78\% for bridge-to-transplant
\cite{patel2014contemporary}. Therefore, there is great need for
reliable predictions of patient-specific risk to experience adverse
events after receiving MCS. This information can be used to improve
patient selection for MCS, inform the design of next generation pumps,
and refine patient management strategies.

The Interagency Registry for Mechanically Assisted Circulatory Support
(INTERMACS) was launched to improve MCS patient outcomes through the
collection of patient characteristics, medical events, and long terms
outcomes at a nation wide level. Currently, INTERMACS comprises over
20,000 patients who have received a MCS device. As the largest registry
for data on patients receiving MCS devices, INTERMACS has been leveraged
to develop numerous risk prediction models for mortality and other types
of adverse events that may occur after receiving a device
\cite{kirklin2017eighth, kormos2019society, Adamo950}. A critical step
for developing risk prediction models is handling missing data, which is
common in a ``real world'' database that is not designed for the rigor
of data completeness found in a clinical trial. While some missing
values are related to data entry error, others may relate to instability
of a patient's circulatory status at the time of implant. For instance,
hemodynamic laboratory values are missing for patients in whom invasive
catheterizations were not performed. Patients without invasive
catheterizations may be too ill (unstable) to tolerate the procedure.
Instead, such patients may, after determination of basic hemodynamics,
proceed directly to device placement.

The primary aim of this article is to quantify how much the prognostic
value of a risk prediction model developed from the INTERMACS registry
depends on the strategy that was applied to impute missing data prior to
developing the model. A secondary aim is to measure imputation accuracy
of each strategy by introducing varying levels of artificial missing
data and then imputing it. Because imputation to the mean has been a
standard method for multiple annual summaries of the INTERMACS data, we
measure the potential improvement in prognostic value of a risk
prediction model when other strategies are applied to impute missing
data \emph{instead of imputation to the mean}. The over-arching aim is
to clarify which imputation strategies are most likely to improve risk
prediction (and by extension, quality of care) for patients who receive
MCS devices. This investigation can directly inform future analyses of
INTERMACS data and provide evidence quantifying the benefit of imputing
missing data with sound methodology.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\label{sec:methods}

In this section, we describe the INTERMACS Registry in Section
\ref{subsec:intermacs} and provide a broad overview of general
objectives for imputation of missing data in the context of statistical
inference and supervised statistical learning in Section
\ref{subsec:inference_and_learning}. We then describe thirteen
strategies to handle missing values (Section \ref{subsec:imputation})
and three strategies to develop risk prediction models after missing
data have been imputed (Section \ref{subsec:modeling}). Next, we
describe how risk prediction models and by extension the imputation
strategy applied before training the model were evaluated in Section
\ref{subsec:evaluation}. We outline an internal validation procedure
that was applied to compare combinations of imputation and modeling
strategies using hierarchical Bayesian models in Section
\ref{subsec:internal}. Last, details on computation and software
packages used for the current analysis are described in Section
\ref{subsec:computing}

\hypertarget{intermacs-registry}{%
\subsection{INTERMACS Registry}\label{intermacs-registry}}

\label{subsec:intermacs}

The INTERMACS data is publicly available on biolincc at
\url{https://biolincc.nhlbi.nih.gov/studies/intermacs/}. INTERMACS is a
North American observational registry for patients receiving MCS devices
that began as a partnership between the National Heart, Lung, and Blood
Institute, US Food and Drug Administration, the Centers for Medicaid and
Medicare Services, industry, and individual hospitals with the mission
of improving MCS outcomes. In 2018, INTERMACS became an official Society
of Thoracic Surgeons database.

The current analysis was conducted using publicly available data
provided by the National Heart, Lung, and Blood Institute. We included a
contemporary cohort of 14,738 patients who received continuous flow LVAD
from 2012-2017. Patient follow-up begins after implantation of a
durable, long term MCS device and continues while the device is in
place. Registry endpoints include death on a device, heart
transplantation, or cessation of support (for recovery and non-recovery
reasons). INTERMACS collects pre-implant patient characteristics,
medical status, laboratory values, and many other variables. Data is
collected at regularly scheduled follow-up as well as during adverse
events such as re-hospitalization. This is a secondary analysis of
de-identified data obtained from the National Heart Lung and Blood
Institute. Primary data collection is approved through University of
Alabama Institutional Review Board and at individual sites.

\hypertarget{statistical-inference-and-learning-with-missing-data}{%
\subsection{Statistical Inference and Learning with Missing
data}\label{statistical-inference-and-learning-with-missing-data}}

\label{subsec:inference_and_learning}

\paragraph{Statistical Inference}

Previous research has thoroughly investigated statistical inference in
the context of missing data. In this setting, analysts often create
multiple imputed datasets, replicate an analysis in each of them, and
then pool their results to obtain valid test statistics for hypothesis
testing \cite{rubin2004multiple}. Imputation strategies that create a
single imputed dataset (\textit{e.g., }imputation to the mean) have been
shown to increase type I errors (\textit{i.e., }rejecting a true null
hypothesis) for inferential statistics by artificially reducing the
variance of observed data and ignoring the uncertainty attributed to
missing values \cite{van2018flexible}. To ensure valid inference,
imputation models should leverage outcome variables to predict missing
predictors \cite{sterne2009multiple}. When very few data are missing,
analysts may apply listwise deletion, \textit{i.e., }removing any
observation with at least one missing value. However, listwise deletion
can easily lead to biased inference \cite{van2020rebutting}.

\paragraph{Statistical Learning}

In the presence of missing data, the goal of supervised statistical
learning is to develop a prediction function for an outcome variable
that accurately generalizes to testing data, which may or may not
contain missing values \cite{hastie2009elements}. Because testing data
may contain missing values, listwise deletion is not a feasible strategy
for statistical learning tasks. In contrast to statistical inference,
strategies that create a single imputed dataset are often used for
statistical learning \cite{kuhn2019feature}. Previous work has
emphasized the importance of imputation strategies with greater
accuracy, suggesting that more accurate imputation strategies lead to
better performance of downstream models (\textit{i.e., }models fitted to
the imputed data) \cite{jerez2010missing}. Others have shown that
imputation to the mean, model-based imputation, and multiple imputation
can provide Bayes-optimal prediction models provided certain assumptions
are true, but these assumptions are difficult to validate in applied
settings \cite{josse2019consistency}.

\paragraph{Using outcomes during imputation}

The outcome variable should be used to impute predictor values for
statistical inference, but using outcome variables for this purpose in
supervised learning can have unintended consequences in model
implementation. For instance, suppose we seek to predict an outcome
\(Y\) for a patient in the clinical setting. If the patient is missing
information for a predictor \(X\) and our model's strategy to impute
missing values of \(X\) leverages the observed value of \(Y\), how
should we impute \(X\)? If we do not already know \(Y\), then we cannot
impute \(X\) and hence cannot generate a valid prediction. On the other
hand, if we already know \(Y\), then we do not need to predict it. Due
to these practical considerations and because INTERMACS is often
leveraged to create prediction models for clinical practice
\cite{thomas2014pre}, we did not leverage outcome variables to impute
missing values of predictors in the current analysis.

\hypertarget{missing-data-strategies}{%
\subsection{Missing Data Strategies}\label{missing-data-strategies}}

\label{subsec:imputation}

\paragraph{Single and multiple imputation}

Several of the imputation methods we considered allowed for creation of
one or many imputed datasets. For downstream models fitted to multiple
sets of imputed data, we applied the modeling technique to each imputed
training set, separately, and then used a pooling technique to generate
a single set of predictions based on multiple imputed testing data.
Specifically, we created 10 imputed training and testing sets, then
applied a modeling procedure to each imputed training set and computed
model predictions on the corresponding imputed testing set, which led to
10 sets of predictions. To aggregate these predictions, we computed the
median for each observation. Informal experiments where the mean was
used instead of the median to aggregate predictions showed little or no
difference in the model's prediction accuracy.

\paragraph{Imputation to the mean}

Imputing data to the mean involves three steps. First, numeric and
nominal variables are identified. Second, for each numeric variable, the
mean is computed and used to impute missing values in the corresponding
variable. Third, for each nominal variable, the mode is computed
(because one cannot compute a mean for categorical variables) and used
to impute missing values in the corresponding variable. The computed
means and modes are then stored for future imputation of testing data.
Because imputation to the mean is frequently used in practice, we use it
as a reference for comparison of all other strategies to impute missing
values,

\paragraph{Bayesian Regression}

Imputation with Bayesian regression draws imputed values from the
posterior distribution of model parameters, accounting for uncertainty
in both model error and estimation \cite{rubin2004multiple}. Multiple
imputation with chained equations (MICE), a well known technique for
generating multiply imputed data
\cite{azur2011multiple, van2018flexible, van2006fully}. In each
iteration of MICE, each specified variable in the dataset is imputed
using the other variables in the dataset. These iterations are run until
convergence criteria have been met.

\paragraph{Predictive Mean Matching (PMM)}

PMM computes predicted values from a pre-specified model that treats one
incomplete column in the data, \(X\), as a dependent variable. For each
missing value in \(X\), PMM identifies a set of candidate donors based
on distance in predicted values of \(X\) \cite{landerman1997empirical}.
One donor is randomly drawn from the candidates, and the observed value
of the donor is taken to replace the missing value. When missing values
were imputed using PMM, we applied MICE to form multiple imputed
datasets. For consistency with other approaches, we also generated a
single imputed dataset using PMM by randomly selecting one of the
multiple datasets imputed.

\paragraph{K-Nearest-Neighbors (KNN)}

KNN imputation identifies \(k\) `similar' observations
(\textit{i.e., }neighbors) for each observation with a missing value in
a given variable \cite{chen2000nearest}. A donor value for the current
missing value is generated by sampling or aggregating the values from
the \(k\) nearest neighbors. In the current analysis, we identified 10
nearest neighbors using Gower's distance \cite{gower}. When imputing a
single dataset using KNN, we aggregated values from 10 nearest neighbors
using the median for numeric variables and the mode for categorical
variables. When imputing multiple datasets using KNN, we aggregated
values from 2, 6, 11, 16, and 20 neighbors, separately, to create 5
imputed datasets.

\paragraph{Hot Deck}

Similar to KNN, hot deck imputation finds \(k\) similar observations for
each observation with a missing value in a given variable
\cite{andridge2010hotdeck}. However, hot deck imputation uses a less
computationally intensive approach, either identifying neighbors at
random or using a subset of variables to find similar observations. When
imputing a single dataset using hot deck imputation, we used a selection
of 5 variables simultaneously to identify nearest neighbors. When
imputing multiple datasets using hot deck imputation, we used a separate
numeric variable for each dataset.

\paragraph{Random forests}

Random forests grow an ensemble of de-correlated decision trees, where
each tree is grown using a bootstrapped replicate of the original
training data
\cite{Breiman2001, hothorn2006survival, strobl2007bias, strobl2008conditional, ishwaran2008random, jaeger2019oblique}.
A particularly helpful feature of random forests is their ability to
estimate testing error by aggregating each decision tree's prediction
error on data outside of their bootstrapped sample
(\textit{i.e., }out-of-bag error). In the current analysis, we conduct
MICE using one random forest to impute each variable, separately. For
each imputed dataset, we allowed random forests to be re-fitted until
out-of-bag error stabilized or a maximum number of iterations was
completed. When imputing a single dataset, we used 250 trees per forest
and a maximum of 10 iterations. When imputing multiple datasets, we used
50 trees per forest and applied PMM using the random forest's predicted
values to impute missing data by sampling one value from a pool of 10
potential donors.

\paragraph{Missingness incorporated as an attribute (MIA)}

MIA is a technique that uses missing status as a predictor rather than
explicitly imputing missing values
\cite{twala2009empirical, ding2010investigation}. MIA adds another
category to nominal variables: ``Missing''. For numeric variables, MIA
creates two columns: one where missing values are imputed with positive
infinity and the other with negative infinity. Since PH models are not
compatible with infinite values, we only use MIA in boosting models.
When a decision tree uses a finite cut-point to split a given numeric
variable, it will assess the cut-point using both the positive and
negative infinite imputed columns, and utilize whichever column provides
the best split of the current data. This procedure translates to sending
all missing values to the left or to the right when forming two new
nodes of the decision tree, using whichever direction results in a
better split.

\paragraph{Assumptions}

Each missing data strategy poses different assumptions regarding the
data and the mechanisms that lead to missing values in the data.
Imputation using Bayesian regression makes the same distributional
assumptions as the Bayesian models that are applied. The
\texttt{miceRanger} algorithm does not make any formal distributional
assumptions, as random forests are non-parametric and can thus handle
skewed and multi-modal data as well as categorical data that are ordinal
or non-ordinal. Hot deck and KNN imputation also do not make
distributional assumptions, but implicitly assume that a missing value
for a given observation can be approximated by aggregating observed
values from the \(k\) most similar observations. PMM makes a similar
implicit assumption, but is slightly more robust to skewed or
multi-modal data because imputed values are sampled directly from
observed ones. MIA operates based on an implicit assumption that
missingness itself is informative. It is difficult to validate these
assumptions in applied settings and also likely that downstream models
will perform poorly if an imputation technique's assumptions are
invalid.

\hypertarget{evaluating-imputation-accuracy}{%
\subsection{Evaluating imputation
accuracy}\label{evaluating-imputation-accuracy}}

\label{subsec:imputation_accuracy}

Imputation accuracy was computed for each numerical and nominal
variable, separately. Numeric variable imputation accuracy was measured
using a re-scaled mean-squared error:
\[1 - \frac{\textrm{MSE}(\textrm{current imputation method})}{\textrm{MSE}(\textrm{imputation to the mean})}.\]
This score is greater than 0 if
\(\textrm{MSE}(\textrm{current imputation method})\) is smaller than
\(\textrm{MSE}(\textrm{imputation to the mean})\), equal to 0 if the two
MSEs are equal, and less than 0 if
\(\textrm{MSE}(\textrm{current imputation method})\) is greater than
\(\textrm{MSE}(\textrm{imputation to the mean})\). Nominal variable
imputation accuracy was measured using a re-scaled classification
accuracy:
\[1 - \frac{\textrm{Classification error}(\textrm{current imputation method})}{\textrm{Classification error}(\textrm{imputation to the mean})}.\]
This score is greater than 0 if the classification error of the current
imputation method is less than that of imputation to the mean, equal to
0 if the two classification errors are equal, and less than 0 of the
classification error of the current imputation method is worse than
imputation to the mean. These numeric and nominal scores are analogous
to the more well known \(R^2\) and Kappa statistics, respectively, but
are modified slightly in the current analysis so that imputation to the
mean will always have a score of 0. This modification makes it easier to
compare the accuracy of each imputation strategy directly with
imputation to the mean.

\hypertarget{risk-prediction-models}{%
\subsection{Risk Prediction Models}\label{risk-prediction-models}}

\label{subsec:modeling}

We applied two modeling strategies after imputing missing values:

\begin{itemize}
\tightlist
\item
  Cox proportional hazards (PH) model with forward stepwise variable
  selection
\item
  Gradient boosted decision trees (hereafter referred to as `boosting')
\end{itemize}

A thorough description of stepwise variable selection and boosting can
be found in Sections 6.1.2 and 8.2.2, respectively, of
\emph{Introduction to Statistical Learning}
\cite{james2013introduction}. Sir David Cox's PH model is one of the
most frequently applied methods for the analysis of right-censored
time-to-event outcomes \cite{kleinbaum2010survival}. According to the PH
assumption, the effect of a unit increase in a predictor is
multiplicative with respect to a baseline hazard function. Boosting
grows a sequence of decision trees, each using information from the
previous trees in an attempt to correct their errors
\cite{friedman2001greedy, chen2016xgboost}.

\hypertarget{evaluation-of-predictions}{%
\subsection{Evaluation of Predictions}\label{evaluation-of-predictions}}

\label{subsec:evaluation}

\paragraph{The Brier score}

The prognostic value of each risk prediction model was primarily
assessed using the Brier score, which depends on both the discrimination
and calibration of predicted risk values
\cite{graf1999assessment, rufibach2010use}. Let \(\tilde{Y}_i(t)\)
represent the observed status of individual \(i\) at time \(t > 0\) in a
testing set of \(M\) observations. Suppose \(\tilde{Y}_i(t)=1\) if there
is an observed event at or before \(t\) and \(\tilde{Y}_i(t)=0\)
otherwise. The Brier score is computed with
\begin{equation} \label{eqn:brier_score}
\widehat{\textrm{BS}}(t)= \frac{1}{M} \sum_{i=1}^{M} \widehat{W}_i(t) \left\{ \tilde{Y}_i(t) - \widehat{S}(t \mid \bm{x}_i) \right\}^2, \end{equation}
where, for the \(i^{th}\) observation, \(\widehat{S}(t \mid \bm{x}_i)\)
is the estimated probability of survival at time \(t\) according to a
given risk prediction model, \(\bm{x}_i\) is the set of input values for
predictor variables in the model, and \(\widehat{W}_i(t)\) is the
inverse proportional censoring weight at time \(t\)
\cite{gerds2006consistent}. Thus, the Brier score is the mean squared
difference between observed event status and expected event status
according to a RPE at time \(t\). Throughout the current analysis, we
set \(t = 6\) months after receiving MCS to focus on short term risk
prediction. Models have been developed to simultaneously predict short
term and long term mortality risk after receiving MCS, but these are
beyond the scope of the current study
\cite{blackstone1986decomposition}.

\paragraph{The scaled Brier score}

The Brier score is dependent on the rate of observed events, which can
make it a difficult metric to interpret. It is often more informative to
scale the Brier score based on the Brier score of a naive model. More
specifically, for a given risk prediction model, the scaled Brier score
is computed as
\[\textrm{Scaled } \widehat{\textrm{BS}}(t)\textrm{ of model } = 1 - \frac{\widehat{\textrm{BS}}(t)\textrm{ of model}}{\widehat{\textrm{BS}}(t)\textrm{ of naive model}}.\]
As the Brier score for risk prediction is analogous to mean-squared
error for prediction of a continuous outcome, the scaled Brier score is
analogous to the \(R^2\) statistic. Similar to the \(R^2\) statistic, a
scaled \(\widehat{\textrm{BS}}(t)\) of 1.00 and 0.00 indicate a perfect
and worthless model, respectively. In our analyses, a Kaplan-Meier
estimate based on the training data (\textit{i.e., }a risk prediction
model that did not use any predictor variables) provided the naive
prediction. In the current analysis, we multiply scaled Brier score
values by 100 for ease of interpretation.

\paragraph{Discrimination and calibration}

The discrimination of a risk prediction model measures the probability
that the model will successfully identify which of two observations is
at higher risk for the event of interest. We estimated discrimination
using a time-dependent concordance (C-) index that accounted for
covariate-dependent censoring. A C-index of 0.50 and 1.00 correspond to
worthless and perfect discrimination, respectively. Similar to the
scaled Brier score, we multiply C-index values by 100 for ease of
interpretation. Calibration slope plots measure a risk prediction
model's absolute accuracy. We estimated calibration error by averaging
the squared distance between expected and observed event rates according
to a calibration plot. Full description of these evaluation metrics are
available \cite{gerds2014calibration, gerds2013estimating}.

\hypertarget{internal-validation-via-monte-carlo-cross-validation-mccv}{%
\subsection{Internal Validation via Monte-Carlo Cross-Validation
(MCCV)}\label{internal-validation-via-monte-carlo-cross-validation-mccv}}

\label{subsec:internal}

To assess the prognostic value of each missing data strategy, we
internally validated a total of 23 modeling algorithm based on
combinations of imputation strategies and modeling strategies described
in Sections \ref{subsec:imputation} and \ref{subsec:modeling}. For
convenience, we use the term `modeling algorithm' to denote the
combination of a missing data strategy and a modeling strategy
(\textit{e.g., }imputation using mean/mode values followed by fitting
the PH model with stepwise variable selection) \cite{kuhn2013applied}.
We conducted internal validation using 200 replicates of Monte-Carlo
cross validation, a resampling technique for internal validation.

\paragraph{Steps taken in each replicate}

In each replicate of Monte-Carlo cross validation, 50\% of the available
data were used for model training and testing. All predictor variables
with \textless{} 50\% missing values were considered for imputation and
subsequent model development. Among these variables, artificial
missingness (0\%, 15\%, or 30\% additional missing values) was induced
based on patient age, with younger and older patients more likely to
have missing data compared to patients who were between 40 and 65 years
of age. Prior to imputation, 50 predictor variables were selected using
a boosting model that quantified variable importance as the fractional
contribution of each predictor to the model based on the total gain
attributed to using the predictor while growing decision trees.
Imputation was conducted in the training and testing sets, separately,
for each imputation strategy. Although some imputation strategies
(\textit{e.g., }KNN and random forests) can impute data in the testing
set using models fitted to the training set, others
(\textit{e.g., }Bayesian regression, PMM, and hot deck) cannot.
Therefore, to ensure fair comparisons in our experiment, each imputation
procedure imputed data in the training set using models fitted to the
training set and then imputed data in the testing set using models
fitted to the testing set. After imputation, Cox PH and boosting models
were applied to each imputed dataset, separately. Last, model
predictions for death and transplant were computed at 6 months following
MCS surgery.

\paragraph{Bayesian analysis of model performance}

To determine whether any of the imputation methods described above had
improved upon imputation to the mean, we applied Bayesian hierarchical
models to analyze differences in scaled Brier score
\cite{benavoli2017time}. This strategy provides a flexible framework to
conduct hypothesis testing and also accounts for correlation occurring
within each replicate of Monte-Carlo cross validation. Specifically,
within each replicate of Monte-Carlo cross validation, the performance
of different modeling algorithms are correlated because they are trained
and tested using the same data.

\hypertarget{statistical-analysis}{%
\subsection{Statistical analysis}\label{statistical-analysis}}

Participant characteristics were tabulated for the overall population
and stratified by event status. Continuous and categorical variables
were summarized as mean (standard deviation) and percent, respectively.
The number and proportion of missing values were also tabulated for the
overall population and stratified by event status. Missing data patterns
were visualized for the overall population using an UpSet plot
\cite{lex2014upset}. All of the proceeding analyses were conducted using
resampling results from Monte-Carlo cross validation. Imputation
accuracy was aggregated for all numeric and nominal variables to create
two overall scores for each imputed dataset. For imputation methods that
created multiple datasets, scores were aggregated over each dataset to
provide a single summary score. The distribution
(\textit{i.e., }25\textsuperscript{th}, 50\textsuperscript{th}, and
75\textsuperscript{th} percentile) of the scaled Brier score was
estimated for each modeling algorithm.

We split our results from Monte-carlo cross validation into four
datasets based on outcome (mortality or transplant) and modeling
procedure (Cox PH or boosting). For each dataset, we fit one
hierarchical Bayesian model where the dependent variable was scaled
Brier score and independent variables included the imputation strategy
applied and the amount of artificially missing data (0\%, 15\%, or 30\%)
induced before imputation. With each model, we estimated the difference
in scaled Brier score of downstream models when random forests, Bayesian
regression, PMM, hot deck, MIA, or KNN imputation were applied to impute
missing values instead of imputation to the mean.

\hypertarget{computational-details}{%
\subsection{Computational details}\label{computational-details}}

\label{subsec:computing}

Our source code is available on GitHub (see
\url{https://github.com/bcjaeger/INTERMACS-missing-data}).

\hypertarget{results}{%
\section{Results}\label{results}}

\paragraph{Patient Characteristics}

During the first 6 months after receiving MCS, the number (\% of overall
population) of deaths, transplants, and loss-to-follow-up events was
1,897 (13\%), 1,067 (7.2\%), and 981 (6.7\%), respectively. The mean
(standard deviation) age of patients was 57 (13) years, 66\% of patients
identified as white and 79\% were male (Table
\ref{tbl_characteristics}).

\paragraph{Missing data}

Many predictor variables exhibited similar proportions of missing values
in different outcome groups (Table \ref{tbl_missingness}). However, the
number (percent) of missing values for surgery time was an exception,
with 6,125 (41.6\%) in the overall population, 2,246 (53.4\%) among
patients who died, and 1,713 (27.2\%) among patients who were censored.
Additionally, missing values for surgery time were frequently
accompanied by missing values for CV pressure (Figure \ref{fig:upset}).

\paragraph{Imputation accuracy}

Single imputation strategies were more accurate than their counterparts
using multiple imputation in 8 out of 10 comparisons of nominal scores
and 10 out of 10 comparisons of numeric scores (Table
\ref{tbl_impute_accuracy}). When an additional 15\% of data were
missing, single imputation KNN obtained the highest nominal accuracy
(score: +0.02, 95\% CI -0.05, 0.15) and single imputation with random
forests obtained the highest numeric accuracy (score: +0.05, 95\% CI
0.03, 0.08). When an additional 30\% of data were missing, imputation to
the mean obtained the highest numeric and nominal scores.

\hypertarget{scaled-brier-score}{%
\subsection{Scaled Brier score}\label{scaled-brier-score}}

\paragraph{Mortality risk prediction}

When no additional data were amputed and imputation to the mean was
applied before fitting risk prediction models, the median
(25\textsuperscript{th}, 75\textsuperscript{th} percentile) scaled Brier
score was 6.09 (5.71, 6.54) for Cox PH and 7.27 (6.83, 7.77) for
boosting models (Table \ref{tbl_md_strat_dead_ipa}; top panel). Multiple
imputation strategies universally obtained higher scaled Brier scores
versus their single imputation counterparts. Multiple imputation using
random forests provided the highest scaled Brier score compared to other
strategies, leading to a median (25\textsuperscript{th},
75\textsuperscript{th} percentile) increase in the scaled Brier score of
0.34 (0.12, 0.54) for Cox PH and 0.34 (0.14, 0.58) for boosting models.
These performance increments improved when an aditional 15\% and 30\% of
data were amputed (Table \ref{tbl_md_strat_dead_ipa}; middle and bottom
panel).

\paragraph{Transplant risk prediction}

When no additional data were amputed and imputation to the mean was
applied before fitting risk prediction models, the median
(25\textsuperscript{th}, 75\textsuperscript{th} percentile) scaled Brier
score was 9.35 (8.80, 9.71) for Cox PH and 9.00 (8.55, 9.41) for
boosting models (Table \ref{tbl_md_strat_txpl_ipa}; top panel). For Cox
PH models, imputation to the mean provided the lowest scaled Brier
score, and random forest imputation led to a median
(25\textsuperscript{th}, 75\textsuperscript{th} percentile) increase of
0.13 (0.03, 0.29) versus imputation to the mean. For boosting models,
surprisingly, imputation to the mean provided a higher scaled Brier
score than all imputation methods except for multiple imputation using
Bayesian regression. When an additional 15\% and 30\% of data were
amputed, the performance increments corresponding to the use of multiple
imputation increased for both Cox PH and boosting models (Table
\ref{tbl_md_strat_txpl_ipa}; middle and bottom panel).

\hypertarget{discrimination-and-calibration}{%
\subsection{Discrimination and
calibration}\label{discrimination-and-calibration}}

\paragraph{Mortality risk prediction}

Regardless of how much additional data were amputed, boosting models
obtained higher median C-indices than PH models for prediction of
6-month mortality risk (Table \ref{tbl_md_strat_dead_auc}). In addition,
all models that used multiple imputation consistently obtained higher
median C-indices compared to their counterparts using single imputation.
When 0 and 15\% additional data were amputed, median calibration error
was lower for PH models compared to boosting models, but boosting models
obtained lower median calibration error when 30\% additional data were
amputed (Table \ref{tbl_md_strat_dead_cal_error}). Almost all imputation
strategies provided lower median calibration error compared with
imputation to the mean.

\paragraph{Transplant risk prediction}

For all amounts of additional data amputed, PH models obtained higher
median C-indices than boosting models for prediction of 6-month
transplant risk (Table \ref{tbl_md_strat_txpl_auc}). Similar to
mortality risk prediction, multiple imputation strategies generally
provided higher C-indices than their counterparts using single
imputation strategy. For boosting models, MIA provided higher C-indices
compared to all other single imputation strategies and had similar or
superior performance compared to several multiple imputation strategies.
Differences in calibration error were minor when no additional missing
data were amputed (Table \ref{tbl_md_strat_txpl_cal_error}). When an
additional 30\% of data were amputed, boosting models using MIA obtained
lower calibration error than any other strategy.

\paragraph{Bayesian analysis of model performance}

Adjusting for the amount of additional missing data amputed and the
outcome variable, the posterior probability that an imputation strategy
would improve the scaled Brier score of a downstream model relative to
imputation to the mean was maximized by using multiple imputation with
random forests (probability of improvement: 0.998; Figure
\ref{fig:fig_md_strat_infer_ipa}). Similarly, multiple imputation using
random forests was estimated to have the highest posterior probability
of improving the C-index in comparison to using imputation to the mean
(probability of improvement: 0.962; Figure
\ref{fig:fig_md_strat_infer_auc}). However, the estimated posterior
probability of a reduction in calibration was 1 when using either single
or multiple imputation with random forests compared with imputation to
the mean (Figure \ref{fig:fig_md_strat_infer_cal_error}). Although
imputation using random forest was estimated to be the best overall
option, there was moderate to strong evidence that MIA and each multiple
imputation strategy we applied would improve prognostic accuracy of
downstream models compared with imputation to the mean.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

In this article, we leveraged data from the INTERMACS registry to
evaluate how the use of different imputation strategies prior to fitting
a risk prediction model would impact the external prognostic accuracy of
the model. External prognostic accuracy was measured at 6 months after
receiving MCS to focus on short term risk prediction, and the primary
measure of accuracy was the scaled Brier score. We evaluated the
performance of 12 imputation strategies in a broad range of settings by
varying the type of modeling algorithm, the outcome variable, and the
amount of additional missing data amputed prior to performing
imputation. Our resampling experiment indicated that conducting multiple
imputation has a high likelihood of increasing the downstream scaled
Brier score and C-index of risk prediction models compared with
imputation to the mean. Additionally, multiple imputation with random
forests emerged as the imputation strategy that maximized the
probability of developing a more prognostic model compared with
imputation to the mean.

In previous studies involving the INTERMACS data registry, imputation to
the mean has been applied prior to developing a mortality risk
prediction model
\cite{hsich2012should, cotts2014predictors, eckman2011survival, kirklin2017eighth, kormos2019society}.
The current analysis has shown that the prognostic accuracy of these
models can be increased by implementing a different strategy to impute
missing values, particularly if multiple imputation is applied. However,
few studies have investigated best practices on the use of multiple
imputation for prediction models, and few software packages allow users
to implement multiple imputation procedures that adhere to best
practices for prediction model development (\textit{e.g., }using only
the training data to impute missing values). Further development of open
source software that allows investigators to safely and succinctly
implement multiple imputation for prediction models (\textit{e.g., }the
\texttt{miceRanger} package) may increase the presence of multiple
imputation in published risk prediction models.

\clearpage
\begin{table}
\caption{Participant characteristics stratified by event status.}
\label{tbl_characteristics}

\begin{tabular}{lccccc}
\toprule
Characteristic & Overall & Censored & Cessation of support & Dead & Transplant\\
\midrule
No. of patients & 14,738 & 6,308 & 310 & 4,204 & 3,916\\
months\_post\_implant & 18 (16) & 25 (18) & 16 (11) & 13 (14) & 13 (10)\\
Age, years & 57 (13) & 57 (13) & 48 (14) & 61 (12) & 53 (12)\\
Sex &  &  &  &  & \\
\hspace{1em}Female & 21 & 22 & 41 & 22 & 19\\
\addlinespace
\hspace{1em}Male & 79 & 78 & 59 & 78 & 81\\
Race &  &  &  &  & \\
\hspace{1em}White & 66 & 63 & 64 & 71 & 67\\
\hspace{1em}Black & 25 & 28 & 27 & 21 & 23\\
\hspace{1em}Other & 9.2 & 9.6 & 9.4 & 7.4 & 10\\
\addlinespace
Body mass index & 29 (7) & 29 (7) & 29 (8) & 29 (7) & 28 (6)\\
Device strategy &  &  &  &  & \\
\hspace{1em}Other & 0.6 & 0.6 & 2.6 & 0.5 & 0.6\\
\hspace{1em}Bridge to transplant & 52 & 43 & 51 & 40 & 81\\
\hspace{1em}Destination therapy & 47 & 56 & 46 & 60 & 18\\
\addlinespace
LVEDD & 6.83 (1.10) & 6.85 (1.07) & 6.49 (0.93) & 6.69 (1.12) & 6.95 (1.12)\\
Urinary albumin, g/dl & 3.41 (0.64) & 3.42 (0.61) & 3.41 (0.64) & 3.33 (0.66) & 3.49 (0.65)\\
Urinary creatinine, mg/dl & 1.39 (0.68) & 1.36 (0.62) & 1.17 (0.53) & 1.49 (0.77) & 1.35 (0.67)\\
CV pressure & 10.8 (6.2) & 10.4 (6.1) & 9.9 (6.0) & 11.6 (6.5) & 10.7 (6.0)\\
Periphal edema & 35 & 33 & 37 & 44 & 31\\
\addlinespace
BUN, mg/dl & 29 (18) & 28 (17) & 22 (13) & 32 (19) & 28 (18)\\
Bilirubin levels, mg/dl & 1.37 (1.84) & 1.25 (1.14) & 1.28 (1.95) & 1.52 (2.46) & 1.41 (1.97)\\
Device type &  &  &  &  & \\
\hspace{1em}Bi-ventricular assistance device & 3.5 & 2.1 & 0.3 & 5.8 & 3.4\\
\hspace{1em}Left-ventricular assistance device & 97 & 98 & 100 & 94 & 97\\
\addlinespace
Surgery time, minutes & 292 (113) & 289 (110) & 272 (96) & 320 (128) & 272 (98)\\
CPB time, minutes & 95 (49) & 93 (45) & 88 (38) & 103 (56) & 90 (44)\\
\bottomrule
\multicolumn{6}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{6}{l}{\rule{0pt}{1em}Table values are mean (standard deviation) or \%}\\
\end{tabular}

\end{table}

\clearpage

\begin{table} 
 \caption{Number (percent) of missing values for a selection of predictor variables in the overall population and in subgroups based on event status.} 
\label{tbl_missingness}

\begin{tabular}{lccccc}
\toprule
 & Overall & Dead & Transplant & Censored & Cessation of support \\ 
\midrule
CV pressure & 7,415 (50.3\%) & 2,156 (51.3\%) & 2,006 (51.2\%) & 3,108 (49.3\%) & 145 (46.8\%) \\ 
Surgery time, minutes & 6,125 (41.6\%) & 2,246 (53.4\%) & 2,007 (51.3\%) & 1,713 (27.2\%) & 159 (51.3\%) \\ 
Periphal edema & 5,093 (34.6\%) & 1,543 (36.7\%) & 1,470 (37.5\%) & 1,952 (30.9\%) & 128 (41.3\%) \\ 
LVEDD & 3,155 (21.4\%) & 1,000 (23.8\%) & 866 (22.1\%) & 1,233 (19.5\%) & 56 (18.1\%) \\ 
Urinary albumin, g/dl & 1,014 (6.88\%) & 326 (7.75\%) & 309 (7.89\%) & 359 (5.69\%) & 20 (6.45\%) \\ 
CPB time, minutes & 1,010 (6.85\%) & 287 (6.83\%) & 274 (7.00\%) & 421 (6.67\%) & 28 (9.03\%) \\ 
Bilirubin levels, mg/dl & 850 (5.77\%) & 291 (6.92\%) & 237 (6.05\%) & 304 (4.82\%) & 18 (5.81\%) \\ 
Body mass index & 77 (0.52\%) & 26 (0.62\%) & 19 (0.49\%) & 31 (0.49\%) & 1 (0.32\%) \\ 
BUN, mg/dl & 50 (0.34\%) & 10 (0.24\%) & 8 (0.20\%) & 32 (0.51\%) & 0 (0.00\%) \\ 
Sex & 24 (0.16\%) & 9 (0.21\%) & 4 (0.10\%) & 11 (0.17\%) & 0 (0.00\%) \\ 
Urinary creatinine, mg/dl & 20 (0.14\%) & 8 (0.19\%) & 4 (0.10\%) & 8 (0.13\%) & 0 (0.00\%) \\ 
Age, years & 0 (0.00\%) & 0 (0.00\%) & 0 (0.00\%) & 0 (0.00\%) & 0 (0.00\%) \\ 
Race & 0 (0.00\%) & 0 (0.00\%) & 0 (0.00\%) & 0 (0.00\%) & 0 (0.00\%) \\ 
Device strategy & 0 (0.00\%) & 0 (0.00\%) & 0 (0.00\%) & 0 (0.00\%) & 0 (0.00\%) \\ 
Device type & 0 (0.00\%) & 0 (0.00\%) & 0 (0.00\%) & 0 (0.00\%) & 0 (0.00\%) \\ 
\bottomrule
\end{tabular} 
 \end{table}

\clearpage

\begin{table} 
 \caption{Accuracy of strategies to impute artificial missing data} 
\label{tbl_impute_accuracy}

\begin{tabular}{lcccc}
\toprule
& \multicolumn{2}{c}{Nominal variables} & \multicolumn{2}{c}{Numeric variables} \\ 
 \cmidrule(lr){2-3}\cmidrule(lr){4-5}
 & Single imputation & Multiple imputation & Single imputation & Multiple imputation \\ 
\midrule
\multicolumn{1}{l}{+15\% additional missing data} \\ 
\midrule
Imputation to the mean & 0 (reference) & --- & 0 (reference) & --- \\ 
Hot deck & -0.24 (-0.34, -0.16) & -0.25 (-0.37, -0.18) & -0.35 (-0.53, -0.29) & -0.42 (-0.54, -0.36) \\ 
K-nearest-neighbors & +0.02 (-0.05, 0.15) & -0.01 (-0.09, 0.11) & +0.01 (-0.01, 0.03) & -0.02 (-0.05, 0.00) \\ 
Predictive mean matching & -0.15 (-0.34, 0.00) & -0.15 (-0.35, 0.00) & -0.25 (-0.40, -0.20) & -0.26 (-0.35, -0.21) \\ 
Random forest & -0.15 (-0.26, -0.03) & -0.15 (-0.26, -0.03) & +0.05 (0.03, 0.08) & +0.02 (-0.01, 0.04) \\ 
Bayesian regression & -0.14 (-0.26, 0.01) & -0.14 (-0.27, 0.01) & -0.30 (-0.41, -0.23) & -0.30 (-0.41, -0.23) \\ 
\midrule
\multicolumn{1}{l}{+30\% additional missing data} \\ 
\midrule
Imputation to the mean & 0 (reference) & --- & 0 (reference) & --- \\ 
Hot deck & -0.25 (-0.34, -0.18) & -0.26 (-0.36, -0.18) & -0.34 (-0.46, -0.30) & -0.40 (-0.53, -0.36) \\ 
K-nearest-neighbors & -0.04 (-0.09, 0.03) & -0.07 (-0.13, 0.00) & -0.02 (-0.03, -0.01) & -0.05 (-0.07, -0.04) \\ 
Predictive mean matching & -0.18 (-0.38, -0.04) & -0.18 (-0.37, -0.04) & -0.29 (-0.46, -0.24) & -0.29 (-0.40, -0.25) \\ 
Random forest & -0.18 (-0.28, -0.07) & -0.18 (-0.28, -0.08) & -0.01 (-0.03, 0.01) & -0.03 (-0.06, -0.01) \\ 
Bayesian regression & -0.17 (-0.29, -0.04) & -0.17 (-0.29, -0.05) & -0.31 (-0.43, -0.27) & -0.31 (-0.43, -0.27) \\ 
\bottomrule
\end{tabular} 
 \end{table}

\clearpage

\begin{table} 
 \caption{Median (25th, 75th percentile) change in scaled Brier score when different imputation strategies are applied to training and testing sets instead of imputation to the mean prior to developing a risk prediction model for mortality. Table values show the scaled Brier score for imputation to the mean. For other imputation strategies, table values show the change in scaled Brier score relative to the scaled Brier score when imputation to the mean was applied. All table values are multiplied by 100 for ease of interpretability} 
\label{tbl_md_strat_dead_ipa}

\begin{tabular}{lcccc}
\toprule
& \multicolumn{2}{c}{Proportional hazards} & \multicolumn{2}{c}{Gradient boosted decision trees} \\ 
 \cmidrule(lr){2-3}\cmidrule(lr){4-5}
Imputation method & Single Imputation & Multiple Imputation & Single Imputation & Multiple Imputation \\ 
\midrule
\multicolumn{1}{l}{No additional missing data} \\ 
\midrule
Imputation to the mean & 6.09 (5.71, 6.54) & -- & 7.27 (6.83, 7.77) & -- \\ 
Missingness as an attribute & -- & -- & +0.09 (-0.16, 0.31) & -- \\ 
Hot deck & -0.25 (-0.47, 0.01) & +0.05 (-0.15, 0.24) & -0.24 (-0.46, 0.03) & +0.25 (0.08, 0.45) \\ 
K-nearest-neighbors & +0.09 (-0.06, 0.23) & +0.18 (0.04, 0.34) & -0.06 (-0.27, 0.18) & +0.28 (0.06, 0.46) \\ 
Predictive mean matching & -0.30 (-0.53, -0.03) & +0.15 (-0.01, 0.37) & -0.22 (-0.43, 0.14) & +0.34 (0.13, 0.54) \\ 
Random forest & +0.05 (-0.24, 0.33) & +0.34 (0.12, 0.54) & -0.08 (-0.32, 0.21) & +0.34 (0.14, 0.58) \\ 
Bayesian regression & -0.35 (-0.61, -0.10) & +0.07 (-0.10, 0.29) & -0.25 (-0.52, 0.05) & +0.29 (0.08, 0.52) \\ 
\midrule
\multicolumn{1}{l}{+15\% additional missing data} \\ 
\midrule
Imputation to the mean & 5.56 (4.95, 6.03) & -- & 6.72 (6.13, 7.33) & -- \\ 
Missingness as an attribute & -- & -- & +0.52 (0.21, 0.77) & -- \\ 
Hot deck & +0.10 (-0.22, 0.46) & +0.49 (0.21, 0.78) & +0.02 (-0.28, 0.33) & +0.58 (0.29, 0.92) \\ 
K-nearest-neighbors & +0.36 (0.10, 0.65) & +0.53 (0.25, 0.79) & +0.04 (-0.23, 0.36) & +0.42 (0.16, 0.68) \\ 
Predictive mean matching & +0.03 (-0.32, 0.41) & +0.54 (0.20, 0.90) & +0.21 (-0.25, 0.64) & +0.66 (0.32, 1.12) \\ 
Random forest & +0.39 (-0.05, 0.75) & +0.76 (0.37, 1.12) & +0.35 (0.01, 0.76) & +0.73 (0.41, 1.06) \\ 
Bayesian regression & +0.05 (-0.44, 0.44) & +0.49 (0.10, 0.85) & +0.19 (-0.21, 0.59) & +0.58 (0.29, 1.06) \\ 
\midrule
\multicolumn{1}{l}{+30\% additional missing data} \\ 
\midrule
Imputation to the mean & 4.79 (4.11, 5.55) & -- & 6.29 (5.67, 7.02) & -- \\ 
Missingness as an attribute & -- & -- & +0.40 (0.11, 0.75) & -- \\ 
Hot deck & +0.47 (0.03, 0.99) & +0.73 (0.22, 1.38) & +0.10 (-0.24, 0.44) & +0.51 (0.05, 0.99) \\ 
K-nearest-neighbors & +0.57 (0.23, 0.97) & +0.77 (0.47, 1.18) & -0.09 (-0.44, 0.23) & +0.36 (0.09, 0.67) \\ 
Predictive mean matching & +0.25 (-0.36, 0.82) & +0.71 (0.16, 1.36) & +0.09 (-0.33, 0.64) & +0.58 (0.06, 1.05) \\ 
Random forest & +0.62 (0.08, 1.22) & +1.04 (0.39, 1.83) & +0.39 (-0.19, 0.77) & +0.69 (0.27, 1.12) \\ 
Bayesian regression & +0.22 (-0.40, 0.82) & +0.68 (0.10, 1.28) & +0.10 (-0.44, 0.62) & +0.43 (-0.03, 0.93) \\ 
\bottomrule
\end{tabular} 
 \end{table}

\clearpage

\begin{table} 
 \caption{Median (25th, 75th percentile) change in scaled Brier score when different imputation strategies are applied to training and testing sets instead of imputation to the mean prior to developing a risk prediction model for transplant. Table values show the scaled Brier score for imputation to the mean. For other imputation strategies, table values show the change in scaled Brier score relative to the scaled Brier score when imputation to the mean was applied. All table values are multiplied by 100 for ease of interpretability} 
\label{tbl_md_strat_txpl_ipa}

\begin{tabular}{lcccc}
\toprule
& \multicolumn{2}{c}{Proportional hazards} & \multicolumn{2}{c}{Gradient boosted decision trees} \\ 
 \cmidrule(lr){2-3}\cmidrule(lr){4-5}
Imputation method & Single Imputation & Multiple Imputation & Single Imputation & Multiple Imputation \\ 
\midrule
\multicolumn{1}{l}{No additional missing data} \\ 
\midrule
Imputation to the mean & 9.35 (8.80, 9.71) & -- & 9.00 (8.55, 9.41) & -- \\ 
Missingness as an attribute & -- & -- & -0.07 (-0.28, 0.13) & -- \\ 
Hot deck & +0.01 (-0.12, 0.15) & +0.06 (-0.04, 0.23) & -0.24 (-0.43, 0.04) & -0.01 (-0.20, 0.17) \\ 
K-nearest-neighbors & +0.06 (-0.04, 0.17) & +0.08 (0.00, 0.19) & -0.27 (-0.50, -0.05) & -0.05 (-0.20, 0.10) \\ 
Predictive mean matching & +0.04 (-0.12, 0.19) & +0.14 (0.02, 0.29) & -0.34 (-0.60, -0.10) & -0.06 (-0.23, 0.15) \\ 
Random forest & +0.03 (-0.10, 0.19) & +0.13 (0.03, 0.29) & -0.32 (-0.50, -0.07) & -0.05 (-0.21, 0.14) \\ 
Bayesian regression & +0.03 (-0.13, 0.16) & +0.10 (0.00, 0.26) & -0.32 (-0.58, -0.05) & +0.02 (-0.13, 0.25) \\ 
\midrule
\multicolumn{1}{l}{+15\% additional missing data} \\ 
\midrule
Imputation to the mean & 8.76 (8.20, 9.21) & -- & 8.01 (7.57, 8.54) & -- \\ 
Missingness as an attribute & -- & -- & +0.50 (0.27, 0.78) & -- \\ 
Hot deck & +0.25 (-0.04, 0.55) & +0.42 (0.20, 0.78) & -0.10 (-0.34, 0.23) & +0.21 (-0.02, 0.52) \\ 
K-nearest-neighbors & +0.46 (0.20, 0.73) & +0.54 (0.28, 0.79) & -0.10 (-0.41, 0.20) & +0.12 (-0.12, 0.38) \\ 
Predictive mean matching & +0.45 (0.10, 0.76) & +0.65 (0.34, 0.94) & -0.03 (-0.34, 0.21) & +0.25 (0.02, 0.62) \\ 
Random forest & +0.44 (0.12, 0.76) & +0.60 (0.32, 0.95) & -0.07 (-0.35, 0.26) & +0.16 (-0.09, 0.47) \\ 
Bayesian regression & +0.53 (0.21, 0.79) & +0.74 (0.37, 0.94) & -0.07 (-0.39, 0.24) & +0.24 (-0.09, 0.56) \\ 
\midrule
\multicolumn{1}{l}{+30\% additional missing data} \\ 
\midrule
Imputation to the mean & 7.96 (7.26, 8.63) & -- & 7.09 (6.45, 7.68) & -- \\ 
Missingness as an attribute & -- & -- & +0.83 (0.55, 1.10) & -- \\ 
Hot deck & +0.45 (-0.02, 0.90) & +0.60 (0.15, 1.11) & -0.01 (-0.34, 0.41) & +0.19 (-0.17, 0.65) \\ 
K-nearest-neighbors & +0.42 (0.00, 0.79) & +0.61 (0.18, 0.90) & -0.21 (-0.54, 0.19) & +0.09 (-0.22, 0.46) \\ 
Predictive mean matching & +0.70 (0.21, 1.21) & +0.90 (0.46, 1.46) & +0.09 (-0.32, 0.54) & +0.34 (-0.04, 0.85) \\ 
Random forest & +0.70 (0.14, 1.27) & +0.93 (0.49, 1.45) & -0.13 (-0.46, 0.43) & +0.25 (-0.12, 0.66) \\ 
Bayesian regression & +0.77 (0.32, 1.40) & +1.00 (0.44, 1.53) & 0.00 (-0.45, 0.52) & +0.24 (-0.17, 0.80) \\ 
\bottomrule
\end{tabular} 
 \end{table}

\clearpage

\begin{table} 
 \caption{Median (25th, 75th percentile) change in concordance index when different imputation strategies are applied to training and testing sets instead of imputation to the mean prior to developing a risk prediction model for mortality. Table values show the concordance index for imputation to the mean. For other imputation strategies, table values show the change in concordance index relative to the concordance index when imputation to the mean was applied. All table values are multiplied by 100 for ease of interpretability} 
\label{tbl_md_strat_dead_auc}

\begin{tabular}{lcccc}
\toprule
& \multicolumn{2}{c}{Proportional hazards} & \multicolumn{2}{c}{Gradient boosted decision trees} \\ 
 \cmidrule(lr){2-3}\cmidrule(lr){4-5}
Imputation method & Single Imputation & Multiple Imputation & Single Imputation & Multiple Imputation \\ 
\midrule
\multicolumn{1}{l}{No additional missing data} \\ 
\midrule
Imputation to the mean & 69.3 (68.9, 69.8) & -- & 70.7 (70.3, 71.0) & -- \\ 
Missingness as an attribute & -- & -- & 0.00 (-0.20, 0.19) & -- \\ 
Hot deck & -0.20 (-0.44, 0.05) & +0.04 (-0.13, 0.27) & -0.19 (-0.36, 0.00) & +0.13 (-0.04, 0.28) \\ 
K-nearest-neighbors & +0.03 (-0.10, 0.23) & +0.14 (-0.02, 0.36) & -0.10 (-0.30, 0.13) & +0.14 (-0.03, 0.30) \\ 
Predictive mean matching & -0.34 (-0.56, -0.03) & +0.11 (-0.05, 0.37) & -0.21 (-0.44, 0.05) & +0.21 (0.06, 0.39) \\ 
Random forest & +0.04 (-0.24, 0.29) & +0.30 (0.08, 0.52) & -0.02 (-0.26, 0.18) & +0.24 (0.07, 0.41) \\ 
Bayesian regression & -0.39 (-0.69, -0.10) & +0.09 (-0.16, 0.28) & -0.21 (-0.44, -0.02) & +0.22 (0.06, 0.41) \\ 
\midrule
\multicolumn{1}{l}{+15\% additional missing data} \\ 
\midrule
Imputation to the mean & 69.2 (68.8, 69.7) & -- & 70.5 (70.1, 71.0) & -- \\ 
Missingness as an attribute & -- & -- & +0.10 (-0.11, 0.31) & -- \\ 
Hot deck & -0.18 (-0.52, 0.19) & +0.25 (-0.07, 0.58) & -0.23 (-0.50, 0.03) & +0.12 (-0.07, 0.32) \\ 
K-nearest-neighbors & +0.16 (-0.17, 0.42) & +0.30 (0.05, 0.56) & -0.16 (-0.33, 0.10) & +0.10 (-0.07, 0.31) \\ 
Predictive mean matching & -0.42 (-0.83, 0.02) & +0.24 (-0.07, 0.53) & -0.26 (-0.54, 0.00) & +0.26 (0.06, 0.43) \\ 
Random forest & -0.04 (-0.39, 0.26) & +0.41 (0.05, 0.64) & -0.08 (-0.29, 0.19) & +0.27 (0.04, 0.43) \\ 
Bayesian regression & -0.50 (-0.95, -0.06) & +0.20 (-0.11, 0.47) & -0.29 (-0.48, -0.03) & +0.23 (0.07, 0.43) \\ 
\midrule
\multicolumn{1}{l}{+30\% additional missing data} \\ 
\midrule
Imputation to the mean & 68.9 (68.4, 69.4) & -- & 70.2 (69.8, 70.7) & -- \\ 
Missingness as an attribute & -- & -- & +0.12 (-0.08, 0.31) & -- \\ 
Hot deck & -0.15 (-0.52, 0.30) & +0.12 (-0.23, 0.51) & -0.22 (-0.47, 0.02) & +0.10 (-0.17, 0.32) \\ 
K-nearest-neighbors & +0.17 (-0.11, 0.49) & +0.33 (0.09, 0.65) & -0.08 (-0.27, 0.16) & +0.19 (-0.01, 0.38) \\ 
Predictive mean matching & -0.75 (-1.15, -0.34) & +0.06 (-0.29, 0.52) & -0.36 (-0.67, -0.04) & +0.27 (0.04, 0.48) \\ 
Random forest & -0.29 (-0.80, 0.23) & +0.26 (-0.10, 0.64) & -0.16 (-0.47, 0.11) & +0.29 (0.07, 0.52) \\ 
Bayesian regression & -0.75 (-1.28, -0.22) & +0.12 (-0.29, 0.48) & -0.36 (-0.70, -0.07) & +0.19 (-0.01, 0.44) \\ 
\bottomrule
\end{tabular} 
 \end{table}

\clearpage

\begin{table} 
 \caption{Median (25th, 75th percentile) change in calibration error when different imputation strategies are applied to training and testing sets instead of imputation to the mean prior to developing a risk prediction model for mortality. Table values show the calibration error for imputation to the mean. For other imputation strategies, table values show the change in calibration error relative to the calibration error when imputation to the mean was applied. All table values are multiplied by 100 for ease of interpretability} 
\label{tbl_md_strat_dead_cal_error}

\begin{tabular}{lcccc}
\toprule
& \multicolumn{2}{c}{Proportional hazards} & \multicolumn{2}{c}{Gradient boosted decision trees} \\ 
 \cmidrule(lr){2-3}\cmidrule(lr){4-5}
Imputation method & Single Imputation & Multiple Imputation & Single Imputation & Multiple Imputation \\ 
\midrule
\multicolumn{1}{l}{No additional missing data} \\ 
\midrule
Imputation to the mean & 2.41 (1.71, 3.44) & -- & 3.29 (2.24, 4.80) & -- \\ 
Missingness as an attribute & -- & -- & -0.09 (-1.07, 0.65) & -- \\ 
Hot deck & -0.04 (-0.68, 0.58) & -0.45 (-1.25, 0.25) & -0.13 (-1.12, 0.65) & -0.90 (-1.98, 0.01) \\ 
K-nearest-neighbors & +0.04 (-0.59, 0.65) & -0.29 (-0.80, 0.39) & -0.30 (-1.14, 0.61) & -0.75 (-1.59, 0.14) \\ 
Predictive mean matching & -0.18 (-0.69, 0.47) & -0.69 (-1.36, -0.03) & -0.11 (-1.27, 0.89) & -1.02 (-2.08, -0.12) \\ 
Random forest & -0.07 (-0.58, 0.65) & -0.46 (-1.10, 0.17) & -0.56 (-1.38, 0.46) & -1.10 (-2.08, -0.14) \\ 
Bayesian regression & -0.08 (-0.75, 0.53) & -0.57 (-1.43, 0.02) & -0.26 (-1.08, 0.57) & -0.85 (-1.91, 0.12) \\ 
\midrule
\multicolumn{1}{l}{+15\% additional missing data} \\ 
\midrule
Imputation to the mean & 5.75 (3.74, 8.59) & -- & 6.28 (4.17, 9.45) & -- \\ 
Missingness as an attribute & -- & -- & -2.83 (-4.81, -0.83) & -- \\ 
Hot deck & -2.16 (-4.31, -1.08) & -3.92 (-6.23, -1.72) & -1.66 (-3.82, -0.34) & -3.44 (-5.97, -1.48) \\ 
K-nearest-neighbors & -2.01 (-3.51, -0.66) & -2.48 (-4.08, -1.14) & -0.61 (-2.28, 0.84) & -2.04 (-3.32, -0.63) \\ 
Predictive mean matching & -3.83 (-6.40, -1.87) & -3.76 (-6.66, -1.04) & -3.52 (-6.43, -1.67) & -3.50 (-7.09, -1.00) \\ 
Random forest & -3.80 (-5.75, -2.07) & -4.35 (-6.60, -1.53) & -3.74 (-6.34, -1.34) & -3.92 (-6.29, -1.50) \\ 
Bayesian regression & -3.76 (-6.45, -1.81) & -3.80 (-6.61, -1.04) & -3.79 (-6.80, -1.39) & -3.27 (-6.40, -0.46) \\ 
\midrule
\multicolumn{1}{l}{+30\% additional missing data} \\ 
\midrule
Imputation to the mean & 9.48 (5.50, 15.3) & -- & 7.02 (4.32, 11.3) & -- \\ 
Missingness as an attribute & -- & -- & -1.61 (-5.29, 0.68) & -- \\ 
Hot deck & -4.95 (-9.18, -2.49) & -6.69 (-12.8, -1.56) & -2.48 (-5.46, -0.31) & -2.48 (-7.18, 0.58) \\ 
K-nearest-neighbors & -3.47 (-6.42, -1.53) & -4.49 (-7.49, -2.39) & +0.73 (-1.56, 3.27) & -1.47 (-3.59, 0.59) \\ 
Predictive mean matching & -7.33 (-13.0, -2.61) & -6.09 (-12.3, -0.22) & -2.58 (-7.90, 0.10) & -1.30 (-6.86, 2.29) \\ 
Random forest & -7.81 (-13.1, -3.63) & -7.39 (-13.4, -2.37) & -3.53 (-7.87, -0.29) & -2.58 (-7.27, 0.44) \\ 
Bayesian regression & -7.27 (-13.2, -2.44) & -5.91 (-12.8, -0.47) & -3.00 (-7.60, 0.97) & -0.54 (-6.60, 3.08) \\ 
\bottomrule
\end{tabular} 
 \end{table}

\clearpage

\begin{table} 
 \caption{Median (25th, 75th percentile) change in concordance index when different imputation strategies are applied to training and testing sets instead of imputation to the mean prior to developing a risk prediction model for transplant. Table values show the concordance index for imputation to the mean. For other imputation strategies, table values show the change in concordance index relative to the concordance index when imputation to the mean was applied. All table values are multiplied by 100 for ease of interpretability} 
\label{tbl_md_strat_txpl_auc}

\begin{tabular}{lcccc}
\toprule
& \multicolumn{2}{c}{Proportional hazards} & \multicolumn{2}{c}{Gradient boosted decision trees} \\ 
 \cmidrule(lr){2-3}\cmidrule(lr){4-5}
Imputation method & Single Imputation & Multiple Imputation & Single Imputation & Multiple Imputation \\ 
\midrule
\multicolumn{1}{l}{No additional missing data} \\ 
\midrule
Imputation to the mean & 79.4 (79.0, 79.8) & -- & 79.2 (78.9, 79.7) & -- \\ 
Missingness as an attribute & -- & -- & -0.03 (-0.22, 0.14) & -- \\ 
Hot deck & +0.01 (-0.12, 0.17) & +0.07 (-0.05, 0.23) & -0.24 (-0.40, -0.02) & -0.04 (-0.20, 0.15) \\ 
K-nearest-neighbors & +0.08 (-0.02, 0.16) & +0.08 (0.01, 0.17) & -0.21 (-0.40, -0.04) & -0.04 (-0.18, 0.10) \\ 
Predictive mean matching & +0.03 (-0.11, 0.16) & +0.13 (0.01, 0.24) & -0.34 (-0.57, -0.12) & -0.06 (-0.19, 0.16) \\ 
Random forest & +0.03 (-0.09, 0.17) & +0.15 (0.06, 0.24) & -0.26 (-0.46, -0.08) & -0.02 (-0.16, 0.14) \\ 
Bayesian regression & +0.01 (-0.15, 0.16) & +0.10 (0.01, 0.22) & -0.32 (-0.54, -0.10) & +0.02 (-0.13, 0.22) \\ 
\midrule
\multicolumn{1}{l}{+15\% additional missing data} \\ 
\midrule
Imputation to the mean & 78.9 (78.5, 79.4) & -- & 78.4 (77.9, 78.9) & -- \\ 
Missingness as an attribute & -- & -- & +0.47 (0.26, 0.71) & -- \\ 
Hot deck & +0.24 (0.04, 0.51) & +0.47 (0.26, 0.76) & -0.13 (-0.40, 0.22) & +0.29 (0.07, 0.58) \\ 
K-nearest-neighbors & +0.36 (0.21, 0.54) & +0.46 (0.29, 0.62) & -0.13 (-0.39, 0.17) & +0.17 (-0.06, 0.39) \\ 
Predictive mean matching & +0.32 (0.11, 0.58) & +0.56 (0.37, 0.82) & -0.12 (-0.38, 0.25) & +0.32 (0.14, 0.61) \\ 
Random forest & +0.32 (0.13, 0.64) & +0.59 (0.36, 0.82) & -0.11 (-0.40, 0.24) & +0.22 (-0.03, 0.51) \\ 
Bayesian regression & +0.39 (0.17, 0.63) & +0.60 (0.40, 0.83) & -0.15 (-0.46, 0.27) & +0.27 (-0.01, 0.64) \\ 
\midrule
\multicolumn{1}{l}{+30\% additional missing data} \\ 
\midrule
Imputation to the mean & 78.4 (77.7, 78.8) & -- & 77.4 (76.8, 78.0) & -- \\ 
Missingness as an attribute & -- & -- & +0.91 (0.52, 1.24) & -- \\ 
Hot deck & +0.34 (-0.03, 0.72) & +0.74 (0.48, 1.11) & -0.19 (-0.53, 0.34) & +0.55 (0.19, 0.86) \\ 
K-nearest-neighbors & +0.28 (-0.02, 0.47) & +0.43 (0.21, 0.66) & -0.32 (-0.78, -0.04) & -0.01 (-0.33, 0.31) \\ 
Predictive mean matching & +0.62 (0.29, 0.98) & +1.02 (0.74, 1.34) & +0.03 (-0.31, 0.44) & +0.65 (0.29, 0.95) \\ 
Random forest & +0.59 (0.22, 1.00) & +1.04 (0.76, 1.30) & -0.21 (-0.58, 0.15) & +0.36 (0.01, 0.68) \\ 
Bayesian regression & +0.81 (0.41, 1.08) & +1.10 (0.80, 1.39) & -0.04 (-0.54, 0.37) & +0.47 (0.10, 0.91) \\ 
\bottomrule
\end{tabular} 
 \end{table}

\clearpage

\begin{table} 
 \caption{Median (25th, 75th percentile) change in calibration error when different imputation strategies are applied to training and testing sets instead of imputation to the mean prior to developing a risk prediction model for transplant. Table values show the calibration error for imputation to the mean. For other imputation strategies, table values show the change in calibration error relative to the calibration error when imputation to the mean was applied. All table values are multiplied by 100 for ease of interpretability} 
\label{tbl_md_strat_txpl_cal_error}

\begin{tabular}{lcccc}
\toprule
& \multicolumn{2}{c}{Proportional hazards} & \multicolumn{2}{c}{Gradient boosted decision trees} \\ 
 \cmidrule(lr){2-3}\cmidrule(lr){4-5}
Imputation method & Single Imputation & Multiple Imputation & Single Imputation & Multiple Imputation \\ 
\midrule
\multicolumn{1}{l}{No additional missing data} \\ 
\midrule
Imputation to the mean & 1.36 (0.92, 2.07) & -- & 1.79 (1.34, 2.50) & -- \\ 
Missingness as an attribute & -- & -- & +0.01 (-0.53, 0.54) & -- \\ 
Hot deck & -0.01 (-0.37, 0.29) & 0.00 (-0.39, 0.26) & +0.11 (-0.56, 0.68) & -0.07 (-0.67, 0.67) \\ 
K-nearest-neighbors & +0.00 (-0.25, 0.24) & -0.02 (-0.25, 0.21) & +0.09 (-0.40, 0.61) & -0.03 (-0.52, 0.67) \\ 
Predictive mean matching & -0.04 (-0.34, 0.32) & -0.08 (-0.41, 0.26) & +0.09 (-0.49, 0.63) & -0.04 (-0.63, 0.70) \\ 
Random forest & -0.06 (-0.39, 0.29) & -0.06 (-0.35, 0.29) & -0.02 (-0.49, 0.63) & +0.02 (-0.54, 0.56) \\ 
Bayesian regression & -0.01 (-0.40, 0.27) & -0.13 (-0.38, 0.18) & +0.12 (-0.46, 0.78) & +0.24 (-0.46, 1.04) \\ 
\midrule
\multicolumn{1}{l}{+15\% additional missing data} \\ 
\midrule
Imputation to the mean & 2.48 (1.79, 3.93) & -- & 2.51 (1.81, 3.29) & -- \\ 
Missingness as an attribute & -- & -- & -0.50 (-1.38, 0.16) & -- \\ 
Hot deck & -0.76 (-1.95, 0.21) & -0.51 (-2.14, 0.89) & -0.01 (-1.00, 0.66) & +0.69 (-0.47, 1.97) \\ 
K-nearest-neighbors & -0.82 (-2.01, 0.21) & -0.78 (-1.96, 0.26) & +0.20 (-0.59, 1.01) & +0.23 (-0.73, 1.02) \\ 
Predictive mean matching & -1.01 (-2.15, 0.27) & -0.46 (-2.16, 0.94) & +0.15 (-0.91, 1.02) & +0.72 (-0.81, 1.87) \\ 
Random forest & -0.80 (-2.32, 0.51) & -0.59 (-2.03, 0.81) & -0.10 (-1.15, 0.66) & +0.42 (-0.93, 1.34) \\ 
Bayesian regression & -0.94 (-2.32, 0.58) & -0.64 (-2.04, 0.89) & +0.19 (-0.76, 1.16) & +0.99 (-0.50, 2.27) \\ 
\midrule
\multicolumn{1}{l}{+30\% additional missing data} \\ 
\midrule
Imputation to the mean & 4.98 (3.42, 7.00) & -- & 3.99 (2.74, 6.22) & -- \\ 
Missingness as an attribute & -- & -- & -1.52 (-2.62, -0.35) & -- \\ 
Hot deck & -2.17 (-4.65, 0.10) & -0.49 (-3.74, 2.34) & -0.53 (-2.79, 1.05) & +1.89 (-0.93, 4.76) \\ 
K-nearest-neighbors & -1.88 (-3.34, -0.22) & -1.88 (-3.70, -0.14) & -0.41 (-1.72, 0.68) & -0.54 (-2.26, 0.88) \\ 
Predictive mean matching & -1.30 (-4.52, 1.28) & +0.04 (-3.48, 2.70) & -0.05 (-2.54, 2.53) & +1.36 (-1.26, 4.01) \\ 
Random forest & -1.80 (-4.34, 0.79) & -0.58 (-3.95, 2.18) & -0.55 (-2.64, 1.53) & +0.33 (-2.27, 2.77) \\ 
Bayesian regression & -1.34 (-4.22, 1.59) & -0.23 (-3.15, 2.61) & -0.16 (-2.47, 1.94) & +0.89 (-1.89, 3.67) \\ 
\bottomrule
\end{tabular} 
 \end{table}

\clearpage

\begin{figure}

{\centering \includegraphics{doc_arxiv_files/figure-latex/upset-1}

}

\caption{An upset plot showing three variables from the INTERMACS registry and all combinations of missing patterns. The bottom left plot shows the number of missing values for each variable, separately. The top right plot shows the number of missing values for each combination of the three variables. For example, there were 2,618 rows in the overall INTERMACS data where both CV pressure and surgery time were missing.}\label{fig:upset}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics{doc_arxiv_files/figure-latex/fig_md_strat_infer_ipa-1}

}

\caption{Posterior distribution of differences in scaled Brier score values relative to imputation to the mean when different imputation strategies are applied before fitting a risk prediction model. Results are aggregated over scenarios where the outcome is mortality and transplant and the amount of additional missing data is 0\%, 15\%, or 30\%. Posterior probability that the difference in scaled Brier score exceeds 0, indicating an improvement in overall model accuracy for the current imputation strategy, is printed to the right of each distribution.}\label{fig:fig_md_strat_infer_ipa}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics{doc_arxiv_files/figure-latex/fig_md_strat_infer_auc-1}

}

\caption{Posterior distribution of differences in scaled Brier score values relative to imputation to the mean when different imputation strategies are applied before fitting a risk prediction model. Results are aggregated over scenarios where the outcome is mortality and transplant and the amount of additional missing data is 0\%, 15\%, or 30\%. Posterior probability that the difference in scaled Brier score exceeds 0, indicating an improvement in overall model accuracy for the current imputation strategy, is printed to the right of each distribution.}\label{fig:fig_md_strat_infer_auc}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics{doc_arxiv_files/figure-latex/fig_md_strat_infer_cal_error-1}

}

\caption{Posterior distribution of differences in calibration error values relative to imputation to the mean when different imputation strategies are applied before fitting a risk prediction model. Results are aggregated over scenarios where the outcome is mortality and transplant and the amount of additional missing data is 0\%, 15\%, or 30\%. Posterior probability that the difference in calibration error exceeds 0, indicating an improvement in overall model accuracy for the current imputation strategy, is printed to the right of each distribution.}\label{fig:fig_md_strat_infer_cal_error}
\end{figure}

\clearpage

\bibliographystyle{unsrt}
\bibliography{references.bib}

\end{document}