### 2020-11-10 README.Rmd

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# INTERMACS-missing-data

[![DOI](https://zenodo.org/badge/298893426.svg)](https://zenodo.org/badge/latestdoi/298893426)

This repository contains all of the code used in the paper _Improving Outcome Predictions for Patients Receiving Mechanical Circulatory Support by Optimizing Imputation of Missing Values_.