### 2020-11-19 README.Rmd
feedback from RC

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# INTERMACS-missing-data

[![DOI](https://zenodo.org/badge/298893426.svg)](https://zenodo.org/badge/latestdoi/298893426)

This repository contains all of the code used in the paper _Improving Outcome Predictions for Patients Receiving Mechanical Circulatory Support by Optimizing Imputation of Missing Values_.

# ABSTRACT

Risk prediction models play an important role in clinical decision making. When developing risk prediction models, practitioners often impute missing values to the mean. We evaluated the impact of applying other strategies to impute missing values on the prognostic accuracy of downstream risk prediction models, i.e., models fitted to the imputed data. A secondary objective was to compare the accuracy of imputation methods based on artificially induced missing values. To complete these objectives, we used data from the Interagency Registry for Mechanically Assisted Circulatory Support (INTERMACS).

## Methods and Results

We applied twelve imputation strategies in combination with two different modeling strategies for mortality and transplant risk prediction following surgery to receive mechanical circulatory support. Model performance was evaluated using Monte-Carlo cross validation and measured based on outcomes 6-months following surgery using the scaled Brier score, concordance index, and calibration error. We used Bayesian hierarchical models to compare model performance. Multiple imputation with random forests emerged as a robust strategy to impute missing values, increasing model concordance by 0.003 (25th, 75th percentile: 0.0008, 0.052) compared with imputation to the mean for mortality risk prediction using a downstream proportional hazards model. The posterior probability that single and multiple imputation using random forests would improve concordance versus mean imputation was 0.464 and >0.999, respectively.

## Conclusion

Selecting an optimal strategy to impute missing values and applying multiple imputation can improve the prognostic accuracy of downstream risk prediction models.